Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists

### 2024-07-19 paper.Rmd
Edits to paper, including additions by Dough

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
author:
  - name: Ronald M. Visser
    orcid_id: 0000-0001-6966-1729
    roles: [writing, editing, analysis, visualization]
    affiliation: Saxion University of Applied Sciences
  - name: Karsten Lambers 
    roles: [writing, editing]
    orcid_id: 0000-0001-6432-0925
    affiliation: Leiden University
  - name: Kenneth Aitchison 
    roles: [writing, editing, software, project administration]
    orcid_id: 0000-0001-6594-1408
    affiliation: Landward Research Teoranta
  - name: Annemarie Jutte
    orcid_id: 0009-0004-9322-0674
    roles: [writing, editing, software]
    affiliation: Saxion University of Applied Sciences
  - name: Doug Rocks-Macqueen 
    orcid_id: 0000
    roles: [writing, editing, software]
    affiliation: Landward Research Teoranta
  - name: Laura van der Knaap 
    roles: [writing, editing, software]
    orcid_id: 0000
    affiliation: Leiden University
  - name: Iza Romanowska 
    orcid_id: 0000-0002-9487-2111
    roles: [writing, editing]
    affiliation: Aarhus University
  - name: Tom Brughmans
    roles: [writing, editing]
    orcid_id: 0000-0002-1589-7768
    affiliation: Aarhus University
  - name: Adam Pažout
    roles: [writing, editing]
    orcid_id: 0000
    affiliation: Aarhus University
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  distill::distill_article:
    number_sections: yes
    fig_caption: yes
  bookdown::pdf_document2:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
  bookdown::word_document2:
    number_sections: yes
    fig_caption: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
# read data
pre_surveys <- readRDS("data/pre_workshop_surveys.rds")
post_surveys <- readRDS("data/post_workshop_surveys.rds")
# factore levels for charts
factor_levels <- c("Excellent", "Good", "Sufficient", "Insufficient", "Not shared")

@@

-   Development of a hands-on vocational training programme, aligned with the recently published *Agent-Based Modelling for Archaeology: Simulating the Complexity of Societies* [@romanowska2021a].

-   Conversion of the training programme into modular learning resources that are interactive, based on HTML and JavaScript standards.

-   Creation of support materials to encourage use of the OERs.

-   Creation of how-to guides to help trainers incorporate OERs into their teaching

-   Creation of a code and learning materials repository that will allow the project to continue sustainability after the European Union support has ended and to facilitate the creation of an open-source community to manage and improve the OERs in the future.

-   Development of the framework for skills acquisition aligned with the Digital Skills Passport [DOUG: reference]?

-   Promotion through a series of multiplier events.

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team. Recognising the difficulty of delivering practical training in an online environment during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics. The participants of the events were asked to respond to two surveys: one before the start of the event and one completed after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix ... for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix ... for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was given at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

| Conference or event          | Period                  | Online, in person or hybrid | Number of participants | Number of registrations |
|---------------|---------------|---------------|---------------|---------------|
| CAA Amsterdam                | April 2023              | In person                   | 36                     | 53                      |
| EAA Belfast                  | August 2023             | In person                   | 21                     | No registration         |
| CAA-DE/NL-Fl Online workshop | October 2023            | Online                      | 40                     | 76                      |
| Reuvensdagen                 | November 2023           | In person                   | 13                     | No registration         |
| CAA-UK                       | November 2023           | Hybrid                      | 53                     |                         |
| Leiden (course)              | December 2023           | In person                   | 11                     | No registration         |
| Aarhus                       | January/February 2024   | Online                      | 176                    | \>500                   |
| Saxion                       | March 2023-January 2024 | In person                   | 18                     | No registration         |
| **Total**                    |                         |                             | **368**                | **\>628**               |

@@

```{r nationality, fig.cap="The nationality of the participants that filled in the survey.", echo=FALSE}
ggplot(pre_surveys, aes(x = `What.is.you.nationality....Nationality...Text`)) +
  geom_bar() +  coord_flip() + xlab("Nationality")

@@

```{r gender-age, fig.cap="The gender and age distribution of the respondents for each workshop", fig.width=11, echo=FALSE}
ggplot(pre_surveys, aes(x = `What.is.your.age.`, fill = `What.is.your.gender....Selected.Choice`)) +
  geom_bar() + labs(x= "Age", y="Count", fill = "Gender") + facet_grid(~Workshop) +   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r, abm-knowledge, fig.cap="Respondents knowledge and experience with ABM.", fig.show="hold", out.width="50%", echo=FALSE}
pre_surveys %>% 
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>% 
  ggplot(aes(x = `Do.you.know.what.Agent.Based.Modeling.is.`)) +
    geom_bar(na.rm = TRUE) + xlab("I know what ABM is") + facet_grid(~Workshop)
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  ggplot(aes(x=`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`)) + geom_bar() +
  xlab("ABM Software known") + coord_flip()
```

@@

```{r, available-theory, fig.cap="Respondents optionion on the quality of theory on ABM faceted out by event.", fig.width=11, echo=FALSE}
pre_surveys %>%
  select(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`, Workshop) %>%
  filter(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`!="") %>%
  mutate(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.` =
           fct_relevel(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`,
                       "No opinion", "Limited", "Sufficient", "More than sufficient", "Very good")) %>%
  ggplot(aes(x = `How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`)) +
  geom_bar() + xlab("Quality of theory on ABM") + coord_flip() +
  facet_grid(~Workshop)
```

@@

```{r, rating-workshop, fig.cap="Respondents rating of the workshop in general faceted for each event.", fig.width=11, echo=FALSE}
post_surveys %>%
  mutate(How.would.you.rate.the.workshop.in.general. = fct_relevel(How.would.you.rate.the.workshop.in.general., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.workshop.in.general.)) + geom_bar() +
  xlab("Rating of the workshop") + facet_grid(~Workshop) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r, rating-teaching, fig.cap="Respondents rating of the teaching material in general faceted for each event.", fig.width=11, fig.show='hold', out.height="50%", echo=FALSE}
post_surveys %>%
  mutate(How.would.you.rate.the.teaching.material. = fct_relevel(How.would.you.rate.the.teaching.material., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.teaching.material.)) + geom_bar() +
  xlab("Rating of the teaching material") + facet_grid(~Workshop) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
post_surveys %>%
  mutate(How.would.you.rate.the.trainers.teachers. = fct_relevel(How.would.you.rate.the.trainers.teachers., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.trainers.teachers.)) + geom_bar() +
  xlab("Rating of the teacher") + facet_grid(~Workshop) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

## The Digital Competence Framework for Citizens (2.2)

@@

## Digital Skills Passport

>>>>

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
author:
  - name: Ronald M. Visser
    orcid_id: 0000-0001-6966-1729
    roles: [writing, editing, analysis, visualization]
    affiliation: Saxion University of Applied Sciences
  - name: Karsten Lambers 
    roles: [writing, editing]
    orcid_id: 0000-0001-6432-0925
    affiliation: Leiden University
  - name: Kenneth Aitchison 
    roles: [writing, editing, software, project administration]
    orcid_id: 0000-0001-6594-1408
    affiliation: Landward Research Teoranta
  - name: Annemarie Jutte
    orcid_id: 0009-0004-9322-0674
    roles: [writing, editing, software]
    affiliation: Saxion University of Applied Sciences
  - name: Doug Rocks-Macqueen 
    orcid_id: 0000-0001-9021-0260
    roles: [writing, editing, software]
    affiliation: Landward Research Teoranta
  - name: Laura van der Knaap 
    roles: [writing, editing, software]
    orcid_id: 0000
    affiliation: Leiden University
  - name: Iza Romanowska 
    orcid_id: 0000-0002-9487-2111
    roles: [writing, editing]
    affiliation: Aarhus University
  - name: Tom Brughmans
    roles: [writing, editing]
    orcid_id: 0000-0002-1589-7768
    affiliation: Aarhus University
  - name: Adam Pažout
    roles: [writing, editing]
    orcid_id: 0000-0001-7745-5634
    affiliation: Aarhus University
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  distill::distill_article:
    number_sections: yes
    fig_caption: yes
  bookdown::pdf_document2:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
  bookdown::word_document2:
    number_sections: yes
    fig_caption: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
# read data
pre_surveys <- readRDS("data/pre_workshop_surveys.rds")
post_surveys <- readRDS("data/post_workshop_surveys.rds")
# factore levels for charts
factor_levels <- c("Excellent", "Good", "Sufficient", "Insufficient", "Not shared")
# readable event names
event_names <- as_labeller(c("202304_CAA" = "CAA", "202308_EAA" = "EAA", "202310_CAA-DE-NLFl" = "CAA-DE-NLFl", "202311_Reuvensdagen" = "Reuvensdagen", "202311_CAA_UK" = "CAA_UK", "202312_Leiden" = "Leiden", "202401_Aarhus" = "Aarhus"))

@@

-   development of a hands-on vocational training programme, aligned with the recently published *Agent-Based Modelling for Archaeology: Simulating the Complexity of Societies* [@romanowska2021a];

-   conversion of the training programme into modular learning resources that are interactive, based on HTML and JavaScript standards;

-   creation of support materials to encourage use of the OERs;

-   creation of how-to guides to help trainers incorporate OERs into their teaching;

-   creation of a code and learning materials repository that will allow the project to continue sustainability after the European Union support has ended and to facilitate the creation of an open-source community to manage and improve the OERs in the future;

-   development of the framework for skills acquisition aligned with the Digital Skills Passport;

-   promotion through a series of multiplier events.

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team. Recognising the difficulty of delivering practical training in an online environment during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics. The participants of the events were asked to respond to two surveys: one before the start of the event and one completed after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix 1 for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix 2 for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was given at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

| Conference or event          | Period                  | Online, in person or hybrid | Number of participants | Number of registrations |
|---------------|---------------|---------------|---------------|---------------|
| CAA Amsterdam                | April 2023              | In person                   | 36                     | 53                      |
| EAA Belfast                  | August 2023             | In person                   | 21                     | No registration         |
| CAA-DE/NL-Fl Online workshop | October 2023            | Online                      | 40                     | 76                      |
| Reuvensdagen                 | November 2023           | In person                   | 13                     | No registration         |
| CAA-UK                       | November 2023           | Hybrid                      | 53                     | No registration         |
| Leiden (course)              | December 2023           | In person                   | 11                     | No registration         |
| Aarhus                       | January/February 2024   | Online                      | 176                    | \>500                   |
| Saxion                       | March 2023-January 2024 | In person                   | 18                     | No registration         |
| **Total**                    |                         |                             | **368**                | **\>628**               |

@@

```{r nationality, fig.cap="The nationality of the participants that filled in the survey.", fig.height=5, echo=FALSE}
ggplot(pre_surveys, aes(x = `What.is.you.nationality....Nationality...Text`)) +
  geom_bar() +  coord_flip() + xlab("Nationality") + scale_x_discrete(limits=rev)

@@

```{r gender-age, fig.cap="The gender and age distribution of the respondents for each workshop", fig.width=11, echo=FALSE}
ggplot(pre_surveys, aes(x = `What.is.your.age.`, fill = `What.is.your.gender....Selected.Choice`)) +
  geom_bar() + labs(x= "Age", y="Count", fill = "Gender") + facet_grid(~Workshop, labeller = event_names) +   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r, abm-knowledge, fig.cap="Respondents knowledge and experience with ABM.", fig.show="hold", out.width="50%", echo=FALSE}
pre_surveys %>% 
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>% 
  ggplot(aes(x = `Do.you.know.what.Agent.Based.Modeling.is.`)) +
    geom_bar(na.rm = TRUE) + xlab("I know what ABM is") + facet_grid(~Workshop, labeller = event_names)
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  ggplot(aes(x=`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`)) + geom_bar() +
  xlab("ABM Software known") + coord_flip()
```

@@

```{r, available-theory, fig.cap="Respondents optionion on the quality of theory on ABM faceted out by event.", fig.width=11, echo=FALSE}
pre_surveys %>%
  select(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`, Workshop) %>%
  filter(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`!="") %>%
  mutate(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.` =
           fct_relevel(`How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`,
                       "No opinion", "Limited", "Sufficient", "More than sufficient", "Very good")) %>%
  ggplot(aes(x = `How.do.you.judge.the.available.theory.on.Agent.Based.Modelling.`)) +
  geom_bar() + xlab("Quality of theory on ABM") + coord_flip() +
  facet_grid(~Workshop, labeller = event_names)
```

@@

```{r, rating-workshop, fig.cap="Respondents rating of the workshop in general faceted for each event.", fig.width=11, echo=FALSE}
post_surveys %>%
  mutate(How.would.you.rate.the.workshop.in.general. = fct_relevel(How.would.you.rate.the.workshop.in.general., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.workshop.in.general.)) + geom_bar() +
  xlab("Rating of the workshop") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r, rating-teaching, fig.cap="Respondents rating of the teaching material in general faceted for each event.", fig.width=11, fig.show='hold', out.height="50%", echo=FALSE}
post_surveys %>%
  mutate(How.would.you.rate.the.teaching.material. = fct_relevel(How.would.you.rate.the.teaching.material., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.teaching.material.)) + geom_bar() +
  xlab("Rating of the teaching material") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
post_surveys %>%
  mutate(How.would.you.rate.the.trainers.teachers. = fct_relevel(How.would.you.rate.the.trainers.teachers., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.trainers.teachers.)) + geom_bar() +
  xlab("Rating of the teacher") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

## Integration with Learning and Skills Tracking Frameworks

Beyond the creation of the OERs the project also worked on aligning the learning outcomes of the OERs with the Digital Competence Framework for Citizens and the Digital Archaeological Skills Passport.

### The Digital Competence Framework for Citizens (2.2)

@@

### Digital Skills Passport

An archaeological skills passport is a tool for tracking and verifying the archaeological skills that an individual has. The first passport was created as part of *The Ardnamurchan Transitions Project*, a long running investigation of the archaeology and history of the Ardnamurchan Peninsula, Western Scotland, run by the Universities of Leicester and Manchester, with Archaeology Scotland [@cobb2020]. Since then the passport has spread around the world to countries such as Germany [@karl2019] and the United States. While there are many different versions, in its simplest form it is a log booklet listing skills that others can verify by signing off on their level of competencies.

A digital skills passport is the same concept as a physical skills passport but it is digitised to overcome some of the drawbacks of a physical skills passport:

-   losing a copy means losing a record;

-   inability to easily share the log;

-   costs of production;

-   ability to verify those signing off on skill levels.

A trans-Atlantic consortium has been established to create a digital archaeological skills passport. It consists of the professional body in the United States of America, the *Register of Professional Archaeologists* and five archaeological companies in Europe, working and based in Ireland, the United Kingdom, the Netherlands, Germany and France. This is the passport that the ABMA integrate the OERs with.

Agent Based Modelling (ABM) will be placed within the transferable skills category of the Digital Archaeological Skills Passport. This will be facilitated by the creation of a larger category of Analytical Skills and then a sub-category of modelling:

```         
38	Analytical Skills
  38.1 Modelling
	  38.1.1 Computational Modelling
	    38.1.1.1 Agent Based Modelling
	      38.1.1.1.1 Software mastery
	        38.1.1.1.1.1 Creation of models and understanding of key components
	         38.1.1.1.1.2 Integration of other fields - GIS, network science
	      38.1.1.1.2 Analysis and interpretation of results
	 38.1.2 Mathematical Modelling
etc.
```

The digital skills passport has not launched yet, but when it is, it will be found at this url: <https://archaeologypassport.com>.

### 2024-07-19 paper.tex
Edits to paper, including additions by Dough

\title{Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists}
\author{true \and true \and true \and true \and true \and true \and true \and true \and true}
\date{May 22, 2024}

@@

\begin{itemize}
\item
  Development of a hands-on vocational training programme, aligned with the recently published \emph{Agent-Based Modelling for Archaeology: Simulating the Complexity of Societies} (Romanowska, Wren and Crabtree 2021).
\item
  Conversion of the training programme into modular learning resources that are interactive, based on HTML and JavaScript standards.
\item
  Creation of support materials to encourage use of the OERs.
\item
  Creation of how-to guides to help trainers incorporate OERs into their teaching
\item
  Creation of a code and learning materials repository that will allow the project to continue sustainability after the European Union support has ended and to facilitate the creation of an open-source community to manage and improve the OERs in the future.
\item
  Development of the framework for skills acquisition aligned with the Digital Skills Passport {[}DOUG: reference{]}?
\item
  Promotion through a series of multiplier events.
\end{itemize}

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team. Recognising the difficulty of delivering practical training in an online environment during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics. The participants of the events were asked to respond to two surveys: one before the start of the event and one completed after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix \ldots{} for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix \ldots{} for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was given at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
\caption{Conferences, events and situations were workshops were held at which the tutorials were tested.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Conference or event
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Online, in person or hybrid
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of registrations
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Conference or event
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Online, in person or hybrid
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of registrations
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CAA Amsterdam & April 2023 & In person & 36 & 53 \\
EAA Belfast & August 2023 & In person & 21 & No registration \\
CAA-DE/NL-Fl Online workshop & October 2023 & Online & 40 & 76 \\
Reuvensdagen & November 2023 & In person & 13 & No registration \\
CAA-UK & November 2023 & Hybrid & 53 & \\
Leiden (course) & December 2023 & In person & 11 & No registration \\
Aarhus & January/February 2024 & Online & 176 & \textgreater500 \\
Saxion & March 2023-January 2024 & In person & 18 & No registration \\
\textbf{Total} & & & \textbf{368} & \textbf{\textgreater628} \\
\end{longtable}

@@

\hypertarget{the-digital-competence-framework-for-citizens-2.2}{%
\subsection{The Digital Competence Framework for Citizens (2.2)}\label{the-digital-competence-framework-for-citizens-2.2}}

@@

\hypertarget{digital-skills-passport}{%
\subsection{Digital Skills Passport}\label{digital-skills-passport}}

@@

@@

>>>>

\title{Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists}
\author{true \and true \and true \and true \and true \and true \and true \and true \and true}
\date{July 19, 2024}

@@

\begin{itemize}
\item
  development of a hands-on vocational training programme, aligned with the recently published \emph{Agent-Based Modelling for Archaeology: Simulating the Complexity of Societies} (Romanowska, Wren and Crabtree 2021);
\item
  conversion of the training programme into modular learning resources that are interactive, based on HTML and JavaScript standards;
\item
  creation of support materials to encourage use of the OERs;
\item
  creation of how-to guides to help trainers incorporate OERs into their teaching;
\item
  creation of a code and learning materials repository that will allow the project to continue sustainability after the European Union support has ended and to facilitate the creation of an open-source community to manage and improve the OERs in the future;
\item
  development of the framework for skills acquisition aligned with the Digital Skills Passport;
\item
  promotion through a series of multiplier events.
\end{itemize}

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team. Recognising the difficulty of delivering practical training in an online environment during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics. The participants of the events were asked to respond to two surveys: one before the start of the event and one completed after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix 1 for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix 2 for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was given at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
\caption{Conferences, events and situations were workshops were held at which the tutorials were tested.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Conference or event
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Online, in person or hybrid
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of registrations
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Conference or event
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Online, in person or hybrid
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of registrations
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CAA Amsterdam & April 2023 & In person & 36 & 53 \\
EAA Belfast & August 2023 & In person & 21 & No registration \\
CAA-DE/NL-Fl Online workshop & October 2023 & Online & 40 & 76 \\
Reuvensdagen & November 2023 & In person & 13 & No registration \\
CAA-UK & November 2023 & Hybrid & 53 & No registration \\
Leiden (course) & December 2023 & In person & 11 & No registration \\
Aarhus & January/February 2024 & Online & 176 & \textgreater500 \\
Saxion & March 2023-January 2024 & In person & 18 & No registration \\
\textbf{Total} & & & \textbf{368} & \textbf{\textgreater628} \\
\end{longtable}

@@

\hypertarget{integration-with-learning-and-skills-tracking-frameworks}{%
\subsection{Integration with Learning and Skills Tracking Frameworks}\label{integration-with-learning-and-skills-tracking-frameworks}}

Beyond the creation of the OERs the project also worked on aligning the learning outcomes of the OERs with the Digital Competence Framework for Citizens and the Digital Archaeological Skills Passport.

\hypertarget{the-digital-competence-framework-for-citizens-2.2}{%
\subsubsection{The Digital Competence Framework for Citizens (2.2)}\label{the-digital-competence-framework-for-citizens-2.2}}

@@

\hypertarget{digital-skills-passport}{%
\subsubsection{Digital Skills Passport}\label{digital-skills-passport}}

An archaeological skills passport is a tool for tracking and verifying the archaeological skills that an individual has. The first passport was created as part of \emph{The Ardnamurchan Transitions Project}, a long running investigation of the archaeology and history of the Ardnamurchan Peninsula, Western Scotland, run by the Universities of Leicester and Manchester, with Archaeology Scotland (Cobb and Croucher 2020). Since then the passport has spread around the world to countries such as Germany (Karl et al. 2019) and the United States.~While there are many different versions, in its simplest form it is a log booklet listing skills that others can verify by signing off on their level of competencies.~

A digital skills passport is the same concept as a physical skills passport but it is digitised to overcome some of the drawbacks of a physical skills passport:

\begin{itemize}
\item
  losing a copy means losing a record;
\item
  inability to easily share the log;
\item
  costs of production;
\item
  ability to verify those signing off on skill levels.~
\end{itemize}

A trans-Atlantic consortium has been established to create a digital archaeological skills passport. It consists of the professional body in the United States of America, the \emph{Register of Professional Archaeologists} and five archaeological companies in Europe, working and based in Ireland, the United Kingdom, the Netherlands, Germany and France. This is the passport that the ABMA integrate the OERs with.~

Agent Based Modelling (ABM) will be placed within the transferable skills category of the Digital Archaeological Skills Passport. This will be facilitated by the creation of a larger category of Analytical Skills and then a sub-category of modelling:

\begin{verbatim}
38  Analytical Skills
  38.1 Modelling
      38.1.1 Computational Modelling
        38.1.1.1 Agent Based Modelling
          38.1.1.1.1 Software mastery
            38.1.1.1.1.1 Creation of models and understanding of key components
             38.1.1.1.1.2 Integration of other fields - GIS, network science
          38.1.1.1.2 Analysis and interpretation of results
     38.1.2 Mathematical Modelling
etc.
\end{verbatim}

The digital skills passport has not launched yet, but when it is, it will be found at this url: \url{https://archaeologypassport.com}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-cobb2020}{}}%
Cobb, H and Croucher, K. 2020 \emph{Pedagogy, political economy, and training}. In: Cobb, H and Croucher, K (eds.). Oxford University Press. p. 0. DOI: https://doi.org/\href{https://doi.org/10.1093/oso/9780198784258.003.0004}{10.1093/oso/9780198784258.003.0004}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-karl2019}{}}%
Karl, R, Möller, K, Connolly, D, Trausmuth, T, Krierer, KR, Rocks-Macqueen, D, Aitchison, K, Edeso, M, Pintucci, A and Marciniak, A. 2019 Der archäologische Grabungs-Kompetenz-Pass {\textendash} ein Mittel zur Dokumentation durch Praxis erworbener archäologischer Fertigkeiten. \emph{Archäologische Informationen} 42: 237--249. DOI: https://doi.org/\href{https://doi.org/10.11588/ai.2019.0.69362}{10.11588/ai.2019.0.69362}.
